Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-25 paper/paper.tex
updated table 3 and fig 7
done my pass on step 1

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

\vspace{-1ex}
\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
  \vspace{-2ex}
  \caption{Suricata rule for SYN Flood attack.}
  \label{fig:synflood-example}
\end{figure}
\vspace{-2ex}

@@

Unfortunately, the relevant data for the option may be at the message
payload not the fields. We found empirically that this occurs
frequently in http attacks. Figure~\ref{fig:rule-jboss} shows an
example of a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  The content strategy focuses
on extracting such kinds of options from the payload. It is very
important to note that content options are prevalent in \suri\ rules
and the majority of these rule prescribe defenses against http
attacks. For example, \percRulesWithContent{} of the \numrulessuri{}
rules from the public data set~\cite{emerging-threats-open} we
analyzed contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the histogram of
the number of these options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 for
space. Furthermore, \Fix{xxx\%} of the rules are for
http.\MyComment{Rules without \CodeIn{content} options are an
  exception and half of the rules contain at least two
  \CodeIn{content} options.}  Consequently, hangling content options
for http is very important.

We observed, by inspecting existing rules, that \CodeIn{content}
options typically refer to sequence of characters in the payload. To
find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. Considering the example from
Section~\ref{sec:content-example}, \tname{} produced a total of 23
content options.

It is worth noting that we found empirically that deciding which
reverse engineering strategy to use based on the protocol was
effective. We used the payload strategy for http and the field
strategy for the other protocols.\Mar{discuss flood cases...} At the end of this step,
\tname\ produces a seed rule that captures the malicious traffic by
definition.

@@

\begin{table}[h!]
  \small
  \caption{\label{table:results}Results.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
    \#~Options &
    \#~Plausible Rules &    
    \multicolumn{1}{c}{Rank} \\

@@

    \midrule
    Ping Scan & 5 & 2 & - \\
    SYN Flood & 2 & 2 & - \\
    JBoss JMX & 23 & 381 & 23 \\
    WeBid & 23 & - & -\\    
    ColdFusion & 19 & 39 & 5\\
    .htaccess access & 17 & 2458 & 2\\
    .idq access & 19 & 1111 & 2 \\
    iisadmin access & 17 & 861 & 1 \\
    /system32/ in URI & 20 & 3665 & 2 \\
    Script tag in URI & 22 & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=true,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

\vspace{-1ex}
\begin{figure}[h!]
  \lstinputlisting[language=C,numbers=none,keywords={flags,threshold}]{synflood.suricata}
  \vspace{-2ex}
  \caption{\suri\ rule for SYN Flood attack.}
  \label{fig:synflood-example}
\end{figure}
\vspace{-2ex}

@@

Unfortunately, the relevant data for the option may be at the message
payload; not the fields. We found empirically that this occurs
frequently in http attacks, which are prevalent. For example,
\Fix{XXX\%} of the \numrulessuri{} rules from a public data
set~\cite{emerging-threats-open} we analyzed prescribe defenses
against http attacks. Figure~\ref{fig:rule-jboss} shows an example of
a rule that captures traffic containing the string
\CodeIn{"/HtmlAdaptor"} in the payload.  Content options are very
common in http rules---\percRulesWithContent{} of the rules from the
data set mentioned above contain at least one content
option. Figure~\ref{fig:distribution-contents} shows the histogram of
the number of these options per rule for the data set mentioned
above. We limited the number of content options per rule to 10 in the
histogram for space. \MyComment{Rules without \CodeIn{content} options
  are an exception and half of the rules contain at least two
  \CodeIn{content} options.}  Consequently, handling content options
for http is important. The content strategy focuses on extracting such
kinds of options from the payload.

We found, inspecting existing rules, that \CodeIn{content} options
typically refer to sequence of characters in the payload. To find the
values associated with these options, \tname{} splits the payload of
the malicious message in tokens using natural language delimiters,
such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. Considering the example from
Section~\ref{sec:content-example}, \tname{} produced a total of 23
tokens, considered content option candidates.

It is worth noting that we found empirically that deciding which
reverse engineering strategy to use based on the protocol of the input
(malicious) messages was effective. We used the payload strategy for
http and the field strategy for the other protocols. For flooding
attacks, \tname\ runs the same approach for each message and
identifies common options across all messages. For that case,
\tname\ reports, in addition, a \CodeIn{threshold} option, as
discussed on Section~\ref{sec:rules-and-packets}.  At the end of the
Reverse Engineering step, \tname\ produces a seed rule that captures
the malicious traffic.

@@

\begin{table}[h!]
  \small
  \caption{\label{table:results}Results.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    \multicolumn{1}{c}{\multirow{2}{*}{Name}} &
    \#~Option Cand. &
    \#~Plausible Rules &    
    \multicolumn{1}{c}{Rank} \\

@@

    \midrule
    Ping Scan & 5 & 2 & - \\
    SYN Flood & 2 & 2 & - \\
    JBoss JMX & 23 & 381 & 23 \\
    WeBid & 23 & - & -\\    
    ColdFusion & 19 & 39 & 5\\
    .htaccess access & 17 & 2,458 & 2\\
    .idq access & 19 & 1,111 & 2 \\
    iisadmin access & 17 & 861 & 1 \\
    /system32/ in URI & 20 & 3,665 & 2 \\
    Script tag in URI & 22 & - & - \\
    \bottomrule
  \end{tabular}
\end{table}
